---
title: "MiniProjet"
author: "AMIN Carine - DELOUES Damien"
date: "March 30, 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

We have chosen to do an analysis on the alcholic content by volume using Craft beers Dataset.

```{r}
library(ggplot2);
library(dplyr);

beerRaw <- read.csv(file="beers.csv")
```



After loading the data on R Studio, we have selected the columns "style" and "abv", since these 2 columns are the ones we're going to analyse, then we realised that there are NA values in the "abv" column. So we decided to omit them using the function "na.omit()".

Firstly, we decided to do a "boxplot" to have an overview on our selection.

```{r,fig.height=6,fig.width=4}
 beerRaw %>% select(style, abv) %>% na.omit(beerRaw) -> beer
 boxplot(x=beer$abv)

```

As we can see, we have a high standard deviation which indicates that the data points are spread out over a wider range of values. 
We deduce that we have a wide range of beer styles, probably because there are a lot of small breweries that produce original beer styles to attract customers.

We did a graph using "ggplot".
```{r,fig.height=12,fig.width=6}
 beer %>% group_by(style) %>% summarize(NbBiere=n(),MoyenneAbv=mean(abv)) -> beer2
 
 beer2 %>% ggplot(aes(x=MoyenneAbv, y=style)) + geom_point()  + xlab("Average %") +ylab("Beer style")  + theme(axis.text.x =     element_text(angle = 90 )) +  scale_x_continuous(breaks = seq(0,0.15,by=0.005))
```


On the vertical axis we have the different beers style, and on the horizontal axisa rate of the alcoholic content by volume.
With this graphic, we can see that we have a majority of the style have an alcholic content between 4% and 8.5%.
The most alcoholic one is the "English BarleyWine" with more than 10.5% of alchol, and the loess one is the "Low Alcohol Beer" with almost 0%.
We can conclude that the data are quite spread out.

Sadly, the amount of production is missing in the dataset, so we can not say anything about the market consumption.
It would have benn interesting to have the data, and be able to say if the most consumed beers are the strongest or the lightest.



In a second time we decided to relate the average alcoholic content per beer style with the average size of a beer per style, and see if there is any link between those variables that we can make. 

```{r,}
beerRaw %>% select(style, abv, ounces) %>% na.omit(beerRaw) -> beer3

beer3 %>% group_by(style) %>% summarize(NbBiere=n(), MoyenneAbv=mean(abv), MoyenneOunces=mean(ounces)) -> beer4
 
beer4 %>% ggplot(aes(x=MoyenneOunces, y=MoyenneAbv)) +
          geom_point() + 
          xlab("Average ounces") + 
          ylab("Average abv")  + 
          theme(axis.text.x = element_text(angle = 90 ))
```

We could think that more the style has an alcoholic content, smaller is the size.

But with the graph, we can see that our hypothesis is false. A large part of the data are focus on the same area. Except few of them which are beside.
So we can not conclude any link between the average size of the beer style and his alcoholic content.
