---
title: "Data Manipulation with dplyr"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "February 28, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a demonstration of how dplyr works.

First, we need some data.

```{r}
df <- data.frame(
  name = c("John", "Mary", "Alice", "Peter", "Roger", "Phyllis"),
  age = c(13, 15, 14, 13, 14, 13),
  sex = c("Male", "Female", "Female", "Male", "Male", "Female")
)
head(df);
```

Load the necessary packages:

```{r}
library(dplyr);
library(magrittr);
```

Learning how to use the pipe operator (%>%):

```{r}
df %>% head(n=2);
```

You can chain multiple pipes:

```{r}
df %>% head(n=2) %>% nrow;
```

Now, let's use dplyr verbs, starting with _select()_ to select only one column:

```{r}
df %>% select(name);
```

You can also remove some column using the minus operator:

```{r}
df %>% select(-sex);
```

Now, let's use the _filter()_ verb (let's say we need only those of age 13):

```{r}
df %>% filter(age == 13);
```

If we need to re-order, we can use _arrange()_ (you can pass multiple variables as well):

```{r}
df %>% arrange(age);
```

And to create new columns, we can _mutate()_ (you can use existing variable names):

```{r}
df %>% mutate(birth = 2017 - age);
```

The _summarize()_ verb can be used to reduce data (to know the average age, for instance):

```{r}
df %>% summarize(mean_age = mean(age));
```

The _group_by()_ verb can be used to split data, apply some function and combine results (for example, to know the average age depending on sex):

```{r}
df %>% group_by(sex) %>% summarize(mean_age = mean(age));
```

We can also _group_by()_ by multiple variables (to see the number of occurrences of age/sex combinations):

```{r}
df %>% group_by(age, sex) %>% summarize(N=n());
```

How to use in combination with _ggplot()_:

```{r, fig.width=2.5, fig.height=3}
library(ggplot2);
df %>%
   group_by(sex) %>%
   summarize(mean_age = mean(age)) %>%
   ggplot(aes(x=sex, y=mean_age)) +
          geom_point() + ylim(0,NA) + theme_bw();
```
